Analysing Constraint Grammar with a SAT solver

### 2015-10-11 paper/lrec/problem_desc.tex
more text

\begin{itemize}
\item Is there a sentence that triggers rule(s) X but not rule(s) Y
\item Give a sentence that is ambiguous but doesn't trigger any rules
\item Restrict tag combinations within one word
\end{itemize}

The technique requires an existing morphological lexicon, but no corpus.
This makes it suitable also for languages without large corpora.
Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus. 
Hand-annotated corpora are commonly used in the development of CGs
 \cite{voutilainen99}, because they give immediate feedback whether a new rule
increases or decreases accuracy, and by how large margin.

>>>>

\begin{itemize}
\item Is there a sentence that triggers rule(s) X but not rule(s) Y
\item Give a sentence that is ambiguous but doesn't trigger any rules
\item Does my grammar contain ``dead code'' -- rules that would not fire in any case?
\end{itemize}

The technique requires an existing morphological lexicon, but no corpus.
This makes it suitable also for languages without large corpora.
Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus. 
Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin (\cite{voutilainen2004}). For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe

Things you can get from corpus:

\begin{itemize}
\item \# times the rule was applied
\item \% times it was correct
\end{itemize}
Strengths: gives concrete feedback about the effect of rules. Good especially at the beginning stages of grammar writing.

Things you can get from ML tuning:
\begin{itemize}
\item Variations in the place of the rule
\item Variations in the conditions (careful, ...) of the rule
\end{itemize}
Good side: try all combinations to find the best result.

Things you can get from SAT solver:
\begin{itemize}
\item Interaction of rules; 
\item Create word sequences to test what is possible and what can never happen
\end{itemize}

Given a large morphological lexicon, it is relatively easy to restrict each individual word to contain only realistic ambiguities. We expand the lexicon, map each wordform to its possible analyses, and then ignore the actual wordforms. For instance, a Spanish morphological lexicon contains the entries \texttt{casa:<verb><sg><p3>} and \texttt{casa:<noun><sg>}, hence we know that a confusion between a 3rd person singular verb and a singular noun is possible. The lexicon is not likely to contain a wordform that can be analysed both as an adjective and punctuation, so that kind of ambiguous word is never created.
However, since our method is corpus-free, it does not contain any restrictions as to what words can follow each other. \footnote{Encode morphological lexicon and CG rules in SAT + ask for a sentence that will trigger rules = natural language generation! \:D/}
